Syntax Is All You Need: A Universal-Language Approach to Mutant Generation

### 2023-09-25 paper_fse_acm/sample-acmsmall-conf.tex
added acm formatted version

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,review, anonymous]{acmart}
\acmConference[ICSE 2024]{46th International Conference on Software Engineering}{April 2024}{Lisbon, Portugal}

\usepackage{code}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{listings}
\usepackage{booktabs}
%\usepackage{subfig}
\usepackage{float}
\usepackage{subcaption}
\usepackage{url}
\usepackage{comment}
\usepackage{xcolor}
\usepackage{xspace}

\usepackage{tabularx} % for tables with adjustable column width
\usepackage{makecell} % for formatting table cell contents

@@

%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Syntax Is All You Need: A Universal-Language Approach to Mutant Generation}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
%\author{Alex Groce}
%\affiliation{\institution{Northern Arizona University}\country{United States}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.

%% Table shortcuts
\newcommand{\mr}[2]{\multirow{#1}{*}{#2}}
\newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
\newcommand{\um}{\texttt{universalmutator}\xspace}
%% comments
\newcommand{\clg}[1]{\textcolor{blue}{#1}}
\newcommand{\adg}[1]{\textcolor{purple}{#1}}
\newcommand{\kj}[1]{\textcolor{olive}{#1}}

@@

\begin{document}

\begin{abstract}
While mutation testing has been a topic of academic interest for
decades, it is only recently that ``real-world'' developers, including
industry leaders such as Google and Meta, have adopted mutation
testing.  In this paper we propose a new approach to the development of mutation
testing tools, and in particular the core challenge of
\emph{generating mutants}.  Current practice tends towards two
limited approaches to mutation generation: mutants are either (1)
generated at the bytecode/IR level, and thus neither human readable
nor adaptable to source-level features of languages or projects, or
(2) generated at the source level by language-specific tools that are
hard to write and maintain, and in fact are often abandoned by both
developers and users.  We propose instead that source-level mutation
generation is a special case of \emph{program transformation} in
general, and that adopting this approach allows for a single tool that
can effectively generate source-level mutants for essentially
\emph{any} programming language. Furthermore, by using \emph{parser
  parser combinators} many of the seeming limitations of an
any-language approach can be overcome, without the need to parse
specific languages.  We compare the universal program transformation
approach to mutation to existing tools, and demonstrate the advantages
of using parser parser combinators to improve on a simple regular-expression
based approach to generation.  Finally, we show that our approach
can provide effective mutant generation even for a language for which
it lacks any language-specific operators, and that is not very similar
in syntax to any language it has been applied to previously.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10011001</concept_id>
<concept_desc>Software and its engineering~Dynamic analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011099.10011102.10011103</concept_id>
<concept_desc>Software and its engineering~Software testing and debugging</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

\maketitle

@@

On the positive side, this argues that we can replace a
complex implementation requiring parsing and analysis with
withone that only requires string search-and-replace.
Such an implementation on face can be done using much less code. 
% It also allows for simpler definitions of 
% One can imagine an implementation of the above approach to arithmetic
% operation replacement being implemented in as little as 10-20 lines of
% Python code, including a call to a compiler to check for invalid
%mutants.  
Moreover, defining new mutation operators is not a matter of
understanding a parsed program representation, but is essentially
a matter of providing two strings, where the second is a proposed
replacement for the first.  This simplifies operator specification from
implementation of
tree transformations into a minimal list. 
%Instead of code for arithmetic operator
%replacement, we can imagine the above hypothetical Python program
%reading in a file of thus expressed mutation operators, and
%implementing an unbounded number of operators in a minimalist fashion.

@@

% % example listing, basically use escape inside and then do use \color and \verb
% \begin{lstlisting}[basicstyle=\footnotesize\ttfamily,numbers=none]
% if(strcmp((?\color{dkgreen}\verb|:[1]|?),"(?\color{dkgreen}\verb|:[2]|?)"))
% \end{lstlisting}

\begin{figure}[hbtp]
\centering
\caption{\small Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\end{figure}

% \begin{table}[hbtp]
% \centering
% \caption{Regular Expression Rule Examples}
% \label{tab:rules}
% \resizebox{\columnwidth}{!}{%

%   \begin{tabular}{|l|l|l|}
%     \hline
%     Universal & C-Like & Language-Specific \\
%     \hline
% {\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
% {\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
%                                                                  {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
% {\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
%                  (1==1)} & {\tt any\_of ==> all\_of}\\
% \hline

% \end{tabular}
% }

% \end{table}

% \begin{table}[hbtp]
% \centering
% \caption{Parser Parser Combinator Rule Examples}
% \label{tab:crules}
% \resizebox{\columnwidth}{!}{%

%   \begin{tabular}{|l|l|l|}
%     \hline
%     Universal & C-Like & Language-Specific \\
%     \hline
% {\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
% {\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
%                                                                  {\tt
%                                                                           :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
% {\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
% \hline

% \end{tabular}
% }
% \end{table}

The left-hand-side Table~\ref{tab:rules} shows examples of universal, C-like, and
language-specific rules for a few languages, taken directly from our
implementation of regular-expression based operators (see Section \ref{sec:imp}).  In all cases, the form of a rule is: {\tt regexp
  ==> replacement}, using Python regular expression and replacement
syntax.  Such rules are often non-trivial to write and read; we show
below that using parser parser combinators often simplifies and
clarifies matters.

@@

% ---------------- C++/Java/Rust table UM Only--------------------------
\begin{table}[hbtp]
\centering
\caption{\small Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile.}
\label{tab:table_cpp1javarust}
\begin{tabularx}{\columnwidth}{X|r|rr|rr}

@@

\mc{6}{c}{\textbf{C++}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
ompl & 525 & 2533 & 23.49\% &  2174  & 25.94\%   \\
{\small libgraphqlparser} & 137 & 741 & 21.19\%  & 440 & 29.77\% \\
{\small rethinkdb\_rebirth} &471 & 2246 & 24.27\% & 2009 & 24.49\% \\
xoreos & 732  & 3876  & 62.41\% & 3297 & 69.15\% \\
libxmljs &630 &  3989 & 22.09\% & 4008 & 19.71\% \\[0.5ex]
tiny-diff.-simulator &  195 & 1282 &58.27\% & 1214 & 63.43\% \\

@@

\begin{tabularx}{\columnwidth}{Xllrrr}
                  & &       & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
 \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\

@@

                          & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\

@@

                         & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule

@@

\begin{tabularx}{\columnwidth}{X|r|cr|cr}

@@

\mc{6}{c}{\textbf{Java}} \\\midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
 pebble         & 279 & 1014 & 8.48\%  & 752 & 11.30\%  \\
 spring-hateoas & 397 & 2479 & 10.00\% & 1384 & 12.43\%  \\
 javacc         & 363 & 1888 & 28.70\% & 1276 & 38.71\%  \\
 coffee-gb      & 463 & 1927 & 11.63\% & 798 & 21.80\% \\
 egads          & 252 & 2454 & 25.43\% & 1163 & 51.33\%  \\
 apk-parser     & 324 & 1583 & 22.17\% & 1199 & 27.77\%  \\\bottomrule

@@

\begin{tabularx}{\columnwidth}{Xllrrr}
              & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
\multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\

@@

              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
& \mr{2}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\

@@

              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule

@@

\mc{6}{c}{\textbf{Rust}} \\ \midrule
                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
cargo release    & 346          & 2014   & 42.25\%  & 2401   & 70.39\% \\
passarine        & 376          & 2294   & 9.29\%   & 1691   & 10.23\% \\
typos            & 463          & 1920   & 20.10\%  & 1565   & 23.39\% \\
ord              & 533          & 2858   & 26.66\%  & 2312   & 27.81\% \\
bazuka           & 285          & 1554   & 29.02\%  & 1322   & 30.41\% \\
strum            & 90           & 428     & 11.92\% & 368     & 13.58\% \\\bottomrule

@@

\begin{tabularx}{\columnwidth}{Xllrrr}
               & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
 \multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
 & \mr{2}{Comby} & Mean     & 554.00 &  1055.83 & 29.30\% \\

@@

               &   & Std.Dev. & 592.63 & 508.30 & 21.62\%  \\\bottomrule

@@

@@

\begin{table*}[hbtp]
{\small
    \centering
    \caption{\small Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
    \label{tab:mutationtools}

    \begin{tabular}{|l|l|l|r|}
    \hline
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \hline
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\\cline{2-4}
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ \hline
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\\cline{2-4}
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ \hline
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\\hline
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
        ALL & Our Tool & removed for blinding & 2244 \\\hline
    \end{tabular}}
    \end{table*}

@@

\begin{table}[htbp]
\centering
\caption{\small Java (Our Implementation vs. PIT vs. LittleDarwin).
  Accurately determining if PIT bytecode mutants are equivalent is intractable;
  LittleDarwin produces no equivalent mutants.}
\label{tab:table_java2}
\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule

>>>>

%%
%% This is file `sample-acmsmall-conf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `acmsmall-conf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-acmsmall-conf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[acmsmall,screen,review,anonymous]{acmart}

\usepackage{code}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{listings}
\usepackage{booktabs}
%\usepackage{subfig}
\usepackage{float}
\usepackage{subcaption}
\usepackage{url}
\usepackage{comment}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{cleveref}

\usepackage{tabularx} % for tables with adjustable column width
\usepackage{makecell} % for formatting table cell contents
\usepackage{varwidth} % for achieving window that spans to the widest row

@@

\newcommand{\mr}[2]{\multirow{#1}{*}{#2}}
\newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
\newcommand{\um}{\texttt{universalmutator}\xspace}
%% comments
\newcommand{\clg}[1]{\textcolor{blue}{#1}}
\newcommand{\adg}[1]{\textcolor{purple}{#1}}
\newcommand{\kj}[1]{\textcolor{olive}{#1}}

@@

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[FSE '24]{Foundations of Software Engineering}{July 2024}{Brazil}

%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Syntax Is All You Need: A Universal-Language Approach to Mutant Generation}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  While mutation testing has been a topic of academic interest for
decades, it is only recently that ``real-world'' developers, including
industry leaders such as Google and Meta, have adopted mutation
testing.  In this paper we propose a new approach to the development of mutation
testing tools, and in particular the core challenge of
\emph{generating mutants}.  Current practice tends towards two
limited approaches to mutation generation: mutants are either (1)
generated at the bytecode/IR level, and thus neither human readable
nor adaptable to source-level features of languages or projects, or
(2) generated at the source level by language-specific tools that are
hard to write and maintain, and in fact are often abandoned by both
developers and users.  We propose instead that source-level mutation
generation is a special case of \emph{program transformation} in
general, and that adopting this approach allows for a single tool that
can effectively generate source-level mutants for essentially
\emph{any} programming language. Furthermore, by using \emph{parser
  parser combinators} many of the seeming limitations of an
any-language approach can be overcome, without the need to parse
specific languages.  We compare the universal program transformation
approach to mutation to existing tools, and demonstrate the advantages
of using parser parser combinators to improve on a simple regular-expression
based approach to generation.  Finally, we show that our approach
can provide effective mutant generation even for a language for which
it lacks any language-specific operators, and that is not very similar
in syntax to any language it has been applied to previously.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10011001</concept_id>
<concept_desc>Software and its engineering~Dynamic analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011099.10011102.10011103</concept_id>
<concept_desc>Software and its engineering~Software testing and debugging</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.

%\keywords{Software Testing, Mutants, Mutation Score, Equivalence}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

@@

On the positive side, this argues that we can replace a
complex implementation requiring parsing and analysis with
one that only requires string search-and-replace.
Such an implementation on face can be done using much less code. 
% It also allows for simpler definitions of 
% One can imagine an implementation of the above approach to arithmetic
% operation replacement being implemented in as little as 10-20 lines of
% Python code, including a call to a compiler to check for invalid
%mutants.  
Moreover, defining new mutation operators is not a matter of
understanding a parsed program representation, but is essentially
a matter of providing two strings, where the second is a proposed
replacement for the first.  This simplifies operator specification from
implementation of
tree transformations into a minimal list. 
%Instead of code for arithmetic operator
%replacement, we can imagine the above hypothetical Python program
%reading in a file of thus expressed mutation operators, and
%implementing an unbounded number of operators in a minimalist fashion.

@@

\begin{figure}[hbtp]
\centering
\captionsetup{justification=centering} % Center-align the caption

  \begin{tabular}{l|l}
    \toprule
    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
    \mc{2}{c}{\textbf{Universal}} \\\midrule
    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
\mc{2}{c}{\textbf{C-Like}} \\\midrule
 {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
 {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
 {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
\toprule
\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
 {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
 {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
\bottomrule
\end{tabular}
\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
\label{tab:rules}
\end{figure}

The left-hand-side of Table~\ref{tab:rules} shows examples of universal, C-like, and
language-specific rules for a few languages, taken directly from our
implementation of regular-expression based operators (see Section \ref{sec:imp}).  In all cases, the form of a rule is: {\tt regexp
  ==> replacement}, using Python regular expression and replacement
syntax.  Such rules are often non-trivial to write and read; we show
below that using parser parser combinators often simplifies and
clarifies matters.

@@

%\begin{tabular}{l|r|rr|rr}
% ---------------- C++/Java/Rust table UM Only--------------------------
\begin{table}[hbtp]
\centering
%\captionsetup{justification=centering}
\footnotesize
\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom). LOC describes the project files considered. Muts. is total number of mutants generated for those files. Invalid mutants are those that don't compile.}
\label{tab:table_cpp1javarust}

\begin{varwidth}{.8\linewidth}
\begin{tabularx}{\linewidth}{X|r|rrr|rrr}

@@

\mc{8}{c}{\textbf{C++}} \\\midrule
                 &              & \multicolumn{3}{c|}{\textbf{Regex}} &\multicolumn{3}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}} &  \textbf{Muts.} &  \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\\midrule
ompl & 525 & 2533 & 23.49\% & 0.05 & 2174  & 25.94\% & 0.07  \\
{libgraphqlparser} & 137 & 741 & 21.19\% & 0.90 & 440 & 29.77\% & 0.89\\
{rethinkdb\_rebirth} &471 & 2246 & 24.27\% & 0.12 & 2009 & 24.49\% & 0.10\\
xoreos & 732  & 3876  & 62.41\% & 0.01 & 3297 & 69.15\% & 0.01\\
libxmljs &630 &  3989 & 22.09\% & 0.62 & 4008 & 19.71\% &
0.64 \\[0.5ex]
tiny-diff.-simulator &  195 & 1282 &58.27\% & 0.05 & 1214 & 63.43\% & 0.04\\

@@

\begin{tabularx}{\linewidth}{Xllrrrr}
                  & &       & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} & \textbf{Mut. Score} \\\midrule
 \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% & 0.29\\

@@

                          & & Std.Dev. & 787.62 & 956.31   & 19.48\%  & 0.38\\\cline{2-7}
  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  & 0.29\\

@@

                         & & Std.Dev. & 745.78 & 1058.33 & 21.66\% & 0.37 \\\bottomrule

@@

\begin{tabularx}{\linewidth}{X|r|crr|crr}

@@

\mc{8}{c}{\textbf{Java}} \\\midrule
                 &              & \multicolumn{3}{c|}{\textbf{Regex}} &\multicolumn{3}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}} &  \textbf{Muts.} &  \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\\midrule
 pebble         & 279 & 1014 & 8.48\% & 0.81 & 752 & 11.30\%  & 0.80 \\
 spring-hateoas & 397 & 2479 & 10.00\% & 0.33 & 1384 & 12.43\% & 0.37 \\
 javacc         & 363 & 1888 & 28.70\% & 0.23 & 1276 & 38.71\% & 0.21 \\
 coffee-gb      & 463 & 1927 & 11.63\% & 0.21 & 798 & 21.80\% & 0.23\\
 egads          & 252 & 2454 & 25.43\% & 0.71 & 1163 & 51.33\% & 0.71\\
 apk-parser     & 324 & 1583 & 22.17\% & 0.07 & 1199 & 27.77\% & 0.08 \\\bottomrule

@@

\begin{tabularx}{\linewidth}{Xllrrrr}
              & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} & \textbf{Mut. Score} \\\midrule
\multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% & 0.39\\

@@

              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  & 0.30\\\cline{2-7}
& \mr{2}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% & 0.40\\

@@

              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% & 0.29 \\\bottomrule

@@

\mc{8}{c}{\textbf{Rust}} \\ \midrule
                 &              & \multicolumn{3}{c|}{\textbf{Regex}} &\multicolumn{3}{c}{\textbf{Comby}}  \\
\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{Muts.} &  \textbf{\% Valid} & \makecell{\textbf{Mut.} \\ \textbf{Score}}\\\midrule
cargo release    & 346          & 2014   & 42.25\%  & 0.85
& 2401   & 70.39\% & 0.80\\
passarine        & 376          & 2294   & 9.29\%  & 0.57
 & 1691   & 10.23\% & 0.52\\
typos            & 463          & 1920   & 20.10\% & 0.76  & 1565   & 23.39\% & 0.77\\
ord              & 533          & 2858   & 26.66\%  & 0.73
& 2312   & 27.81\% & 0.83\\
bazuka           & 285          & 1554   & 29.02\% & 0.62 & 1322   & 30.41\% & 0.73\\
strum            & 90           & 428     & 11.92\% & 0.16 & 368    & 13.58\% & 0.22\\\bottomrule

@@

\begin{tabularx}{\linewidth}{Xllrrrr}
               & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} & \textbf{Mut. Score} \\\midrule
 \multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% & 0.62\\
               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  & 0.24\\\cline{2-7}
 & \mr{2}{Comby} & Mean     & 554.00 &  1055.83 & 29.30\% & 0.65\\

@@

               &   & Std.Dev. & 592.63 & 508.30 & 21.62\%  & 0.25\\\bottomrule

@@

\end{varwidth}

@@

\begin{table}[hbtp]
    \small
    \centering
    \caption{\small Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
    \label{tab:mutationtools}

    \begin{tabular}{l|l|l|r}
    %\toprule
    \toprule
    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
    \midrule
    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\
        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ 
    \midrule
    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\
        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ 
    \midrule
        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\
    \midrule
        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\
    \midrule
        ALL & Our Tool & removed for blinding & 2244 \\
    \bottomrule
    \end{tabular}
\end{table}

@@

\begin{table}[htbp]
\centering
%\small
\captionsetup{justification=centering}
\caption{\small Java (Our Implementation vs. PIT vs. LittleDarwin).
  Accurately determining if PIT bytecode mutants are equivalent is intractable;
  LittleDarwin produces no equivalent mutants.}
\label{tab:table_java2}

\begin{varwidth}{.8\columnwidth}
\begin{tabular}{l|rrr|rr|rr}
\toprule
                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule
